---
output:
    html_notebook:
        toc: true
        toc_float: true
editor_options:
  chunk_output_type: inline
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(magrittr)
library(devtools)
```

# `dwctaxon` - package review

## **Reviewer:** [\@collinschwantes](https://github.com/collinschwantes)

### Review Submitted:
**`r cat(sprintf("**Last updated:** %s", Sys.Date()))`**

***

<br>

This report contains documents associated with the review of **rOpenSci** submitted package:

### **`dwctaxon`: ropensci/software-review**  issue [\#574](https://github.com/ropensci/onboarding/issues/574).

<br>

## Package info

**Description:**

Provides functions to create, manipulate, and validate taxonomic 
    data in compliance with Darwin Core standards 
    (Darwin Core "Taxon" class https://dwc.tdwg.org/terms/#taxon).

**Author:** `r 
    c(
      person(given = "Joel H.",
           family = "Nitta",
           role = c("aut", "cre"),
           email = "joelnitta@gmail.com",
           comment = c(ORCID = "0000-0003-4719-7472")),
      person(given = "Wataru",
           family = "Iwasaki",
           role = c("ctb"),
           comment = c(ORCID = "0000-0002-9169-9245"))
    )`

**repo url:** <https://github.com/joelnitta/dwctaxon>

**website url:** <https://joelnitta.github.io/dwctaxon/>

## Review info


#### See [reviewer guidelines](https://ropensci.github.io/dev_guide/reviewerguide.html) for further information on the rOpenSci review process.

**key review checks:**

- Does the code comply with **general principles in the [Mozilla reviewing guide](https://mozillascience.github.io/codeReview/review.html)**?
- Does the package **comply with the [ROpenSci packaging guide](https://ropensci.github.io/dev_guide/building.html)**?
- Are there **improvements** that could be made to the **code style?**
- Is there **code duplication** in the package that should be reduced?
- Are there **user interface improvements** that could be made?
- Are there **performance improvements** that could be made?
- Is the [**documentation**](https://ropensci.github.io/dev_guide/building.html#documentation) (installation instructions/vignettes/examples/demos) **clear and sufficient**?

Please be respectful and kind to the authors in your reviews. The rOpenSci [code of conduct](https://ropensci.github.io/dev_guide/policies.html#code-of-conduct) is mandatory for everyone involved in our review process.

***

### session info


```{r sessionInfo}
sessionInfo()
```


```{r pkg_dir, echo = F}
pkg_dir <- "/Users/collinschwantes/Documents/rOpenSci/reviews/dwctaxon"
```

## Test installation

### test local `dwctaxon` install:

```{r test-local}
install(pkg_dir, dependencies = T, build_vignettes = T)
```

```{r github-rm}
remove.packages("dwctaxon")
```
#### **comments:**

<!-- record comments on local install here -->

No issues with local install.

No issues removing the package.

***

### test install of `dwctaxon` from GitHub with:

```{r test-github}
devtools::install_github("joelnitta/dwctaxon", dependencies = T, build_vignettes = T)
```

#### **comments:**

<!-- record comments on github install here -->

No issues with github install.

***

## Check package integrity

### run checks on `dwctaxon` source:

```{r check-checks}
devtools::check(pkg_dir)
```
#### **comments:**

<!-- record comments on checks here -->

one error for failing test. `dct_add_row` throws an error.

***

### run tests on `dwctaxon` source:

```{r check-tests}
devtools::test(pkg_dir)
```
#### **comments:**

<!-- record comments on tests here -->

Throws 15 errors that all appear to have to do with expect_snapshot

***


### check `dwctaxon` for goodpractice:

```{r test-goodpractice}
goodpractice::gp(pkg_dir)
```
#### **comments:**

<!-- record comments on goodpractice here -->
dct_modify_row_single has a lot of if statements that are nicely labelled in the function. Consider whether or not each of those sections of the function could be split into its own, smaller, easier to evaluate function. 

You could also take advantage of `rlang::empty` to test if something is NA or
NULL. 


***

## Check package metadata files

### inspect

- #### [README](https://github.com/joelnitta/dwctaxon)
- #### [DESCRIPTION](https://github.com/joelnitta/dwctaxon/blob/master/DESCRIPTION)
- #### [NAMESPACE](https://github.com/joelnitta/dwctaxon/blob/master/NAMESPACE)

### spell check

```{r spell-check}
devtools::spell_check(pkg_dir)
```

#### **comments:**

<!-- record comments on metadata files here -->

Metadata files look good!

***

## Check documentation

online documentation: **<https://joelnitta.github.io/dwctaxon/>**

* Is the [documentation](https://ropensci.github.io/dev_guide/building.html#documentation) (installation instructions/vignettes/examples/demos) clear and sufficient?


### test `dwctaxon` function help files:

```{r test-help}
help(package = "dwctaxon")
```

#### **comments:**

<!-- record comments on help files here -->

Help renders as expected.

***

### test `dwctaxon` vignettes:

```{r test-vignettes}
vignette(package = "dwctaxon")
```

#### **comments:**

<!-- record comments on vignettes here -->

Vignettes start with an intro to darwin core then move on to the primary
functions. They are easy to follow and well written. 


***

## Test functionality:

- Are there **user interface improvements** that could be made?

Certain functions create missing columns or provide errors for non-dct terms. Its not always obvious which functions will generate new columns, and which wont. See `dct_fill_col` vs `dct_add_row`

Since there is an "add_row" style function, it may be worth including a drop_row
  


- Are there **performance improvements** that could be made?

Everything ran quickly.

```{r free-style, eval=FALSE}
library("dwctaxon")
library(readxl)
library(tidyverse)
library(dwctaxon)
library(taxize)
library(purrr)
library(tibble)

# load in excel files ----

ids <- read_xlsx(path = "2013 USDA Bee IDs.xlsx")

## convert to long data
id_list <- list()

for(i in 1:length(ids)){
  id_name <- stringr::str_to_lower(colnames(ids)[i])
  id_name_taxa <- stringr::str_remove(id_name, pattern = "\\s*\\(.*\\)[\\.{,3}\\d*]*")
  id_name_sex <- "not provided"
  if(stringr::str_detect(id_name_taxa,pattern = "male|female")){
    id_name_sex <- stringr::str_extract(string = id_name_taxa,pattern = "male|female")
    id_name_taxa <- stringr::str_remove_all(string =id_name_taxa,pattern = sprintf("\\s*%s\\s*",id_name_sex))
    id_name_sex <- stringr::str_to_lower(id_name_sex)
  }
  id_name_identifier_1 <- stringr::str_remove(id_name,pattern = sprintf("%s\\s",id_name_taxa))
  id_name_identifier_2 <- stringr::str_remove(id_name_identifier_1,pattern = id_name_sex)
  id_name_identifier_3 <- stringr::str_remove(id_name_identifier_2,pattern = "\\.{3}\\d+")
  id_name_identifier_no_parens <- stringr::str_remove_all(id_name_identifier_3,pattern = "\\(|\\)")
  ucmc_val <- ids[[i]]
  ucmc_val_no_nas <- ucmc_val[!is.na(ucmc_val)]
  ucmc_val_no_nas_fullucmc <- stringr::str_replace(string = ucmc_val_no_nas,pattern = "^CMC",replacement = "UCMC")
  ucmc_val_no_UCMC <- stringr::str_remove_all(ucmc_val_no_nas_fullucmc,pattern = "UCMC ")

  ucmc_val_no_UCMC_padded <- stringr::str_pad(trimws(ucmc_val_no_UCMC),width = 7,side = "left",pad = "0")


  df <- tibble( "UCMC_NUMBER" = trimws(ucmc_val_no_UCMC_padded),
                "TAXA" = trimws(id_name_taxa),
                "SEX" = trimws(id_name_sex),
                "IDENTIFIED_BY" = trimws(id_name_identifier_no_parens) )

  id_list[[i]] <- df
}

id_df <- do.call(rbind,id_list)



id_df_distinct <- id_df  %>%
  distinct(UCMC_NUMBER,.keep_all = TRUE)


# create dct dataset ----
taxa_data <- id_df_distinct %>%
  distinct(TAXA) %>%
  mutate(TAXA = stringr::str_to_sentence(TAXA)) %>%
  rename("scientificName" = "TAXA")

# check that names are properly formatted
sci_name_check <- dwctaxon::dct_check_sci_name(taxa_data,quiet = FALSE)

## check that auto fill functions work
taxa_data <-taxa_data %>%
  mutate(taxonomicStatus = "accepted")

tsn_ids <- taxize::get_tsn(sci_com = taxa_data$scientificName,searchtype = "scientific")
tax_not_found <- attributes(tsn_ids)$match == "not found"

taxa_data$scientificName[tax_not_found]

tsn_rank <- taxize::classification(tsn_ids,db = "itis")



tsn_rank_most_precise <- purrr::map_dfr(tsn_rank,function(x){
  df <- tibble(name = NA,
               rank = NA,
               id = NA)
  if(is.data.frame(x)){
    df <- tail(x,1)
  }
  return(df)
})



# test vignettes ----
dwc_data <- cbind(taxa_data,tsn_rank_most_precise)

dwc_data_terms <- dwc_data %>%
  rename("UCMC_name" = "scientificName") %>%
  rename("scientificName" = "name",
         "taxonRank" = "rank",
         "taxonID" = "id")

# start validation process

## add a row manually
## add honey bees and cemolobus

## I would like there to be a warning/message that taxon ids are autogenerated

dwc_data_add_row <- dwc_data_terms |>
  dct_add_row(
    scientificName = c("Apis mellifera", "Cemolobus ipomoeae"),
    taxonomicStatus = "accepted",
    taxonRank = "species"
  )


## add a dataframe

data_to_add <- tibble(scientificName = c("Agapostemon texanus","Agapostemon virescens"),
                      taxonomicStatus = "accepted",
                      taxonRank = "species")

dwc_data_add_df <- dwc_data_add_row %>%
  dct_add_row(new_dat = data_to_add)

### modify row

"Unlike add row, or add data frame, modify row rearranges the columns to a
different order. It will not add dct terms that are missing in the original
dataset."

dwc_data_add_df |>
  filter(!is.na(taxonID)) %>%
  dct_modify_row(
    taxonID = "5ac64c9b", # taxonID of Trichomanes crassum Copel.
    tax_status = "synonym",
   # acceptedNameUsageID = "72625a44" # won't add in the column automatically
  )

## add another term and test autocomplete

dwc_data_add_df$acceptedNameUsageID <- NA

dwc_data_mod_row <- dwc_data_add_df |>
  dct_add_row(
    taxonID = "fakeId0099", # fake taxonID
    tax_status = "synonym",
    acceptedNameUsageID = "5ac64c9b", # taxonID of Agapostemon virescens
    scientificName = "Agapostemon inventatus",
    taxonRank = "species"
  )

# check modifying a synonym updates other objects
dwc_data_mod_row_test <- dwc_data_mod_row %>%
  filter(!is.na(taxonID)) %>%
  dct_modify_row(
    taxonID = "5ac64c9b", # taxonID of Agapostemon texanus
    tax_status = "synonym",
    acceptedNameUsageID = "72625a44" # taxonID of Agapostemon virescens
    )

## Check the filling function

"This function does append new columns "
dwc_data_mod_fill_test <- dwc_data_mod_row_test %>%
  dct_fill_col(
    fill_to = "acceptedNameUsage",
    fill_from = "scientificName",
    match_to = "taxonID",
    match_from = "acceptedNameUsageID"
  )

tail(dwc_data_mod_fill_test)


"requires from column to a DWC term"
dwc_data_mod_fill_test_2 <- dwc_data_mod_row_test %>%
  dct_fill_col(
    fill_to = "acceptedNameUsage",
    fill_from = "nothing",
    match_to = "taxonID",
    match_from = "acceptedNameUsageID"
  )

"no warning/error if the from field doesnt exist"
dwc_data_mod_fill_test_3 <- dwc_data_mod_row_test %>%
  dct_fill_col(
    fill_to = "scientificName",
    fill_from = "acceptedNameUsage",
    match_to = "taxonID",
    match_from = "acceptedNameUsageID"
  )
tail(dwc_data_mod_fill_test_2)



## validate

dct_validate(dwc_data_mod_fill_test_3)

"found duplicate entries"
"did not mention non-dct terms"

# with summary param
dct_validate(dwc_data_mod_fill_test_3,on_fail = "summary")
"found bad column name"
"is there a row drop function?"


dwc_data_mod_fill_test_3 %>%
  dct_check_mapping(on_success = "logical")


```

```{r parse-functions}
exports <-ls("package:dwctaxon")
exports
```

<!-- experiment with package functions -->

```{r exp-chunk}

dct_fake <- tibble(taxonID = "111111",taxonomicStatus = "accepted", taxonRank = "species",scientificName = "Inventatum ficticious")

## acceptedNameUsage doesnt exist in the dataset but can be used as a fill_from argument

dct_fake |>
  dct_add_row(taxon_id = "777777",sci_name = "Fechum arribus",tax_status = "accepted") |>
  dct_fill_col(fill_to = "scientificName",fill_from = "acceptedNameUsage",   match_to = "acceptedNameUsageID",match_from = "taxonID")

```

#### **comments:**

<!-- record comments on package experimentation here -->

`dct_fill_col` does not throw an error if the `fill_from` argument is a non-existent column that is a dct_term
***

## Inspect code:

- Does the package **comply with the [ROpenSci packaging guide](https://ropensci.github.io/dev_guide/building.html)**?
    * good [function & variable naming?](https://ropensci.github.io/dev_guide/building.html#function-and-argument-naming)
    * good [dependency management](https://ropensci.github.io/dev_guide/building.html#package-dependencies)?  
- Are there **improvements** that could be made to the [**code style?**](https://ropensci.github.io/dev_guide/building.html#code-style)
- Is there **code duplication** in the package that should be reduced?

```{r inspect-code}
pkgreviewr::pkgreview_print_source("dwctaxon")
```
**\* might not be suitable for large packages with many exported functions**


<br>
<br>

#### **comments:**

<!-- record comments on package source code here -->

Dependencies could include minimum version numbers in the description file

[assertr] appears to have a single function call, is it worth adding the whole package as a dependency? 

## Review test suite:

See guidance on [testing](https://ropensci.github.io/dev_guide/building.html#testing) for further details.

### test coverage

```{r pkg_coverage}
covr::package_coverage(pkg_dir)

```

### inspect [tests](https://github.com/joelnitta/dwctaxon/blob/master/tests/testthat)


#### **comments:**

<!-- record comments on testing suite here -->

`dct_options`  has low coverage. Consider expanding coverage based on the
types of values options can accept. 




***
